---
title: "DRIVE Trial Analysis Code"

output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message=FALSE)
```


#### Posterior function  \
Functions to generate the estimated posterior distribution of the parameter of interest (here is the expected driving time)
```{r posterior distribution function, include=FALSE}
post_dist <-function(plaza=plaza_s1,rainbow=rainbow_s1,stateline=stateline_s1,
                     mu0=15,nu0=0.1, sigma0=5,tau0=5, burn=10000, iter=1000000){
  plaza_matrix <- matrix(c(rep(1,length(plaza)),plaza),ncol = 2)
  rainbow_matrix <- matrix(c(rep(2,length(rainbow)),rainbow),ncol = 2)
  stateline_matrix <- matrix(c(rep(3,length(stateline)),stateline),ncol = 2)
  
  Y<-rbind(plaza_matrix,rainbow_matrix,stateline_matrix)
  
  ### starting value
  m <- length(unique(Y[, 1]))
  n <-sv<-ybar<-rep (NA,m)
  for (j in 1:m){
    ybar[j]<- mean(Y[Y[,1]==j,2])
    sv[j]<-var(Y[Y[,1]==j,2])
    n[j]<-sum(Y[,1]==j)
  }
  
  theta <- ybar
  sigma2 <- mean(sv)
  mu <- mean(theta)
  tau2 <- var(theta) 
  
  ##set up MCMC
  #burn<-1000
  #iter<-2500
  S<-burn+iter
  THETA<-matrix(nrow=S, ncol=m)
  SIGMA2<- matrix(nrow=S,ncol=1)
  
  ###MCMC algorithm
  for (s in 1:S){
    #sample new values of the thetas
    for (j in 1:m){
      #Generate a new theta value from its full conditional use the starting pooled sigma2
      sigma2_temp<-sigma2
      mu_n<-((mu0/tau0^2+n[j]*ybar[j]/sigma2_temp)/(1/tau0^2+n[j]/sigma2_temp))
      tau2_n<-(1/tau0^2+n[j]/sigma2_temp)^(-1)
      theta[j]<-rnorm(1,mu_n,sqrt(tau2_n))
    }
    #sample new value of sigma2
    nun<- nu0+sum(n)
    ss<-nu0*sigma0^2
    for (j in 1:m){ 
      ss <- ss+sum((Y[Y[,1]==j,2]-theta[j])^2)
    }
    sigma2<- 1/rgamma(1,nun/2,ss/2)
    
    #save results
    THETA[s,]<-theta
    SIGMA2[s,]<-sigma2
  }
  
  df_post_mean<-data.frame(plaza=THETA[(burn+1):S,1],
                           rainbow=THETA[(burn+1):S,2],
                           stateline=THETA[(burn+1):S,3])
  post_SIGMA2<-SIGMA2[(burn+1):S,1]
  return(list(post_draw_theta=df_post_mean,post_draw_sigma2=post_SIGMA2))
} 
```

### Interim Analysis function  \
Function to perform the interim analysis 
```{r interim analysis function}
interim_analysis<- function(interim=NA,plaza=plaza_s1,
         rainbow=rainbow_s1,
         stateline=stateline_s1){

post_draws_Thetas<-post_dist(plaza=plaza,rainbow=rainbow,stateline=stateline)

plaza_inttemp<-post_draws_Thetas$post_draw_theta$plaza
rainbow_inttemp<-post_draws_Thetas$post_draw_theta$rainbow
stateline_inttemp<-post_draws_Thetas$post_draw_theta$stateline
variance_inttemp<-post_draws_Thetas$post_draw_sigma2

df_post_all<-data.frame(
  post_dist_plaza=plaza_inttemp,
  post_dist_rainbow=rainbow_inttemp,
  post_dist_stateline=stateline_inttemp,
  post_dist_sigma2=post_draws_Thetas$post_draw_sigma2,
  interim=rep(interim,length(plaza_inttemp))
)


###get the quantiles for the estimates
df_thetas <- data.frame(Parameter=c("theta0","theta1","theta2","sigma","theta1-theta0","theta2-theta0"),
                        Description=c("Mean time for Plaza (Control)",
                                      "Mean time for Rainbow",
                                      "Mean time for State Line",
                                      "Standard deviation for drive time",
                                      "Mean difference between Rainbow and Control",
                                      "Mean difference between State Line and Control"),
                        Estimate=c(mean(plaza_inttemp),
                                   mean(rainbow_inttemp),
                                   mean(stateline_inttemp),
                                   sqrt(mean(variance_inttemp)),
                                   mean(rainbow_inttemp-plaza_inttemp),
                                   mean(stateline_inttemp-plaza_inttemp)),
                        quant0.025=c(quantile(plaza_inttemp,0.025),
                                     quantile(rainbow_inttemp,0.025),
                                     quantile(stateline_inttemp,0.025),
                                     sqrt(quantile(variance_inttemp,0.025)),
                                     quantile(rainbow_inttemp-plaza_inttemp,0.025),
                                     quantile(stateline_inttemp-plaza_inttemp,0.025)),
                        quant0.975=c(quantile(plaza_inttemp,0.975),
                                     quantile(rainbow_inttemp,0.975),
                                     quantile(stateline_inttemp,0.975),
                                     sqrt(quantile(variance_inttemp,0.975)),
                                     quantile(rainbow_inttemp-plaza_inttemp,0.975),
                                     quantile(stateline_inttemp-plaza_inttemp,0.975))
                        ) 

df_quantile<-data.frame(
  quantile=c("2.5%","25%","50%","75%","97.5%"),
      plaza=quantile(plaza_inttemp,c(0.025,0.25,0.50,0.75,0.975)),
      rainbow=quantile(rainbow_inttemp,c(0.025,0.25,0.50,0.75,0.975)),
      stateline=quantile(stateline_inttemp,c(0.025,0.25,0.50,0.75,0.975)),
      variance=quantile(post_draws_Thetas$post_draw_sigma2,c(0.025,0.25,0.50,0.75,0.975)),
      interim=rep(interim,5)
  )

#The quantity of interest: find the route that produce the greatest Pr(Max) probability being Max.Here is the less!!
ProbMax_rainbow<-mean(rainbow_inttemp-plaza_inttemp<stateline_inttemp-plaza_inttemp)
ProbMax_stateline<-mean(stateline_inttemp-plaza_inttemp<rainbow_inttemp-plaza_inttemp)

if(ProbMax_rainbow>ProbMax_stateline){
  theta_dmax<-rainbow_inttemp
}else{theta_dmax<-stateline_inttemp}


###Determine the futility and success
#interim_i=1
Decision<-"Inconclusive"
Decision_Futility <- "No"
Decision_Success <- "No"
Futility_criteria<-max(mean(rainbow_inttemp-plaza_inttemp< -1),mean(stateline_inttemp-plaza_inttemp< -1))
Success_criteria<-mean(theta_dmax < plaza_inttemp)

##generate table for the quantity of interest
df_QOI <- data.frame(Parameter=c("P(theta_d < theta0)",
                                 "Max P(thetad-theta0 < -1"),
                     Description=c("Probability of being better than control for the best route",
                                   "Maximum probability of being better than control by one minute"),
                     Estimate=c(mean(theta_dmax < plaza_inttemp),
                                max(mean(rainbow_inttemp-plaza_inttemp< -1),mean(stateline_inttemp-plaza_inttemp< -1)))
                     )

##Decision Rules

if (Futility_criteria<0.1){
  Decision<-"Futility"
  Decision_Futility <- "Yes"
}else if (Success_criteria>0.998){
  Decision<-"Success"
  Decision_Success<- "Yes"
}else {
  Decision<-"Inconclusive"
  Decision_Futility <- "No"
  Decision_Success<- "No"
}

df_Decision <- data.frame(Parameter=c("Pr(theta_d < theta0>0.998",
                                      "Max Pr(theta_d-theta0 < -1) < 0.1"),
                          Description=c("Success Criteria",
                                        "Futility Criteria"),
                          Estimate=c(Decision_Success,Decision_Futility))

df_sum <- data.frame (
  Decision=Decision,
  Futility_criteria=Futility_criteria,
  Success_criteria=Success_criteria,
  Allocation_rainbow=ProbMax_rainbow,
  Allocation_stateline=ProbMax_stateline,
  plaza_mean=mean(plaza_inttemp),
  rainbow_mean=mean(rainbow_inttemp),
  stateline_mean=mean(stateline_inttemp),
  variance=mean(post_draws_Thetas$post_draw_sigma2),
  plaza_n=length(plaza),
  rainbow_n=length(rainbow),
  stateline_n=length(stateline),
  interim=interim
)


return(list(Quantile=df_quantile,
            Sum_report=df_sum,
            post_dist_draw=df_post_all,
            df_thetas=df_thetas,
            df_QOI=df_QOI,
            df_Decision=df_Decision
            ))
}
```


### RAR Allocation function  \
Functions to calculate the probability Pr(Max) for each route that is used for response adaptive randomization or allocation.
```{r Allocation function}
allocation<- 
  function(interim_result=interim_1st,
           new_ObsNum=8,
           block_size=4,
           fixed_allocation=1){

pal_rainbow<-interim_result$Sum_report$Allocation_rainbow[[1]]
pal_stateline<-interim_result$Sum_report$Allocation_stateline[[1]]

#To avoid assigning drives to a route with a minimal chance of being the best route, any probability less than 0.05 is set to zero at that interim and the resulting probability is reallocated among the remaining routes. In this manner, a route may be temporarily dropped but may be re-introduced if the adaptive randomization probability increases at subsequent interims.
pal_rainbow<-ifelse(pal_rainbow<0.05,0,pal_rainbow) #if the probability is less than 0.05 then drop the arm
pal_stateline<-ifelse(pal_stateline<0.05,0,pal_stateline)

#get the allocation for these 8 participants. If the number of observations is a multiple of its integer factors/block size, then we can allocate 1 to the control and draw three in each block according to the probability of these two routes. If we have extra observations then we need to draw 1 by 1 such as the 33rd observations after the third interim. For example for the 33rd object after the third interim, we need to treat them differently,draw based on the probabilities of all three.

#Initialize a vector to store the allocation probability for all three routes when there are extra observations.
pal_plaza_extra<-NA
pal_rainbow_extra<-NA
pal_stateline_extra <-NA
Subjects <- 1:block_size
if (new_ObsNum%%block_size==0){ #if we have exact number of blocks
block_n=new_ObsNum/block_size
 #the allocation to control is fixed, allocate 1 subject to control for every 4 subjects.
allo_plaza=0
allo_rainbow=0
allo_stateline=0

Allo_order<-data.frame(Subject=vector(),
                       Route=vector(),
                       Block_Num=vector())
for (i in 1:block_n) {
allocation<- sample(x=c("rainbow","stateline"),
       size=block_size-fixed_allocation,
       replace=T,
       prob = c(pal_rainbow,pal_stateline)
       )
allo_plaza <-fixed_allocation + allo_plaza
allo_rainbow <- length(allocation[allocation=="rainbow"]) + allo_rainbow
allo_stateline <- length(allocation[allocation=="stateline"]) + allo_stateline

temp_df<-data.frame(Subject=sample(Subjects),
           Route=c("plaza",allocation),
           Block_Num=i)
temp_df<-temp_df[order(temp_df$Subject, decreasing = FALSE),]#reorder the data
Allo_order <- rbind(Allo_order,temp_df)


}


} else {
  block_n=new_ObsNum%/%block_size #the integer division/factor for the number of blocks. so that we can draw 3 at a time     
  extra_n=new_ObsNum%%block_size #the remainder that we need to draw 1 by 1
  allo_plaza=0
  allo_rainbow=0
  allo_stateline=0
  
  Allo_order<-data.frame(Subject=vector(),
                       Route=vector(),
                       Block_Num=vector())
  
        for (i in 1:block_n){ 
        allocation<- sample(x=c("rainbow","stateline"),
                           size=block_size-fixed_allocation,
                           replace=T,
                           prob = c(pal_rainbow,pal_stateline)
                          )
        allo_plaza <-fixed_allocation + allo_plaza
        allo_rainbow <- length(allocation[allocation=="rainbow"]) + allo_rainbow
        allo_stateline <- length(allocation[allocation=="stateline"]) + allo_stateline  
        
        temp_df<-data.frame(Subject=sample(Subjects),
           Route=c("plaza",allocation),
           Block_Num=i)
       temp_df<-temp_df[order(temp_df$Subject, decreasing = FALSE),]
        Allo_order <- rbind(Allo_order,temp_df)
        }
    #for each extra observations, we need to recalculate the probability  
    pal_plaza_extra=1/block_size 
    pal_rainbow_extra=(1-pal_plaza_extra)*pal_rainbow
    pal_stateline_extra=(1-pal_plaza_extra)*pal_stateline    
    allo_extra<-sample(x=c("plaza","rainbow","stateline"),
                         size=extra_n,
                         replace=T,
                         prob=c(pal_plaza_extra,pal_rainbow_extra,pal_stateline_extra)
                         )
      allo_plaza <- length(allo_extra[allo_extra=="plaza"]) + allo_plaza
      allo_rainbow <- length(allo_extra[allo_extra=="rainbow"]) + allo_rainbow
      allo_stateline <- length(allo_extra[allo_extra=="stateline"]) + allo_stateline 
      
      extra_subjects<-1:extra_n
      
      allo_extra_temp <- data.frame(Subject=sample(extra_subjects),
           Route=allo_extra,
           Block_Num=i+0.5)
      allo_extra_temp<-allo_extra_temp[order(allo_extra_temp$Subject, decreasing = FALSE),]
      Allo_order <- rbind(allo_extra_temp,Allo_order)
}


  df_allocation<-data.frame(routes=c("plaza","rainbow","stateline"),
                          Allocation_Num=c(allo_plaza,allo_rainbow,allo_stateline),
                          Allocation_Prob=c(NA,pal_rainbow,pal_stateline),
                          Allocation_extra_Prob=c(pal_plaza_extra,pal_rainbow_extra,pal_stateline_extra)
                          )
  return (list(df_allocation=df_allocation,Allocation_order=Allo_order))
}

```


### First Interim Analysis  \
```{r First Interim Analysis}
#The functions above should be run before executing the following codes
##Input the collected data
#library(flextable)  
library(rmarkdown)
library(readr)


Main_Data <- read_csv("S:/Biostats/BIO-STAT/BISR/FACTS Working Group/FACTS code/Case Study 22 DRIVE/6.Main Analysis/Blinded Data/Data//DRIVETrial_Initial_Phase.csv")

####Replace with the real data collected 
blinded_route<-unique(Main_Data$route_randomized_blinded)

plaza_s1<-Main_Data$time_diff[Main_Data$route_true_blinded==blinded_route[1]] 
rainbow_s1<-Main_Data$time_diff[Main_Data$route_true_blinded==blinded_route[2]]
stateline_s1<-Main_Data$time_diff[Main_Data$route_true_blinded==blinded_route[3]]

set.seed(1234)  
###First interim
  interim_1st<-interim_analysis(interim = 1,
                                plaza=plaza_s1,
                                rainbow=rainbow_s1,
                                stateline=stateline_s1
                                )

  #flextable(interim_1st$Quantile)
  #paged_table(interim_1st$Sum_report)

paged_table(interim_1st$df_thetas)
paged_table(interim_1st$df_QOI)
paged_table(interim_1st$df_Decision)
  

####RAR
  Allocation_2<-allocation(interim_result=interim_1st,
           new_ObsNum=8,
           block_size=4,
           fixed_allocation=1)
 paged_table(Allocation_2$df_allocation)
 paged_table(Allocation_2$Allocation_order)
 
### Get the randomization list
 Randomization_list <- rep(NA,nrow(Allocation_2$Allocation_order))
 
 for (i in 1:nrow(Allocation_2$Allocation_order)){
   if (Allocation_2$Allocation_order$Route[i]=="plaza"){
   Randomization_list[i]= blinded_route[1]
   } else if (Allocation_2$Allocation_order$Route[i]=="rainbow"){
   Randomization_list[i]= blinded_route[2] 
   } else if (Allocation_2$Allocation_order$Route[i]=="stateline") {
   Randomization_list[i]= blinded_route[3]  
   }
 }


```

### Second Interim Analysis  \
The trial was stopped for futility after the second interim analysis. So, we moved to the final analysis with the 16 drive.
```{r Second Interim Analysis, eval=FALSE}
#The functions above should be run before executing the following codes
##Input the collected data
Main_Data_new <- read_csv("S:/Biostats/BIO-STAT/BISR/FACTS Working Group/FACTS code/Case Study 22 DRIVE/6.Main Analysis/Blinded Data/Data/DRIVETrial_Interim_1.csv")

####Replace with the real data collected 
blinded_route<-unique(Main_Data_new$route_randomized_blinded)

plaza_s1<-Main_Data_new$time_diff[Main_Data_new$route_true_blinded==blinded_route[1]] 
rainbow_s1<-Main_Data_new$time_diff[Main_Data_new$route_true_blinded==blinded_route[2]]
stateline_s1<-Main_Data_new$time_diff[Main_Data_new$route_true_blinded==blinded_route[3]]

  
###Second interim
  interim_2nd<-interim_analysis(interim = 2,
                                plaza=plaza_s1,
                                rainbow=rainbow_s1,
                                stateline=stateline_s1
                                )

  #flextable(interim_2nd$Quantile)
  #paged_table(interim_2nd$Sum_report)

paged_table(interim_2nd$df_thetas)
paged_table(interim_2nd$df_QOI)
paged_table(interim_2nd$df_Decision)

```

### Final Analysis  \
```{r Final Analysis}
#The functions above should be run before executing the following codes
##Input the collected data
Main_Data_new <- read_csv("S:/Biostats/BIO-STAT/BISR/FACTS Working Group/FACTS code/Case Study 22 DRIVE/6.Main Analysis/Blinded Data/Data/DRIVETrial_Interim_1.csv")

####Replace with the real data collected 
blinded_route<-unique(Main_Data_new$route_randomized_blinded)

plaza_s1<-Main_Data_new$time_diff[Main_Data_new$route_true_blinded==blinded_route[1]] 
rainbow_s1<-Main_Data_new$time_diff[Main_Data_new$route_true_blinded==blinded_route[2]]
stateline_s1<-Main_Data_new$time_diff[Main_Data_new$route_true_blinded==blinded_route[3]]
  
### Final Analysis (same as second interim)
  final<-interim_analysis(interim = 2,
                                plaza=plaza_s1,
                                rainbow=rainbow_s1,
                                stateline=stateline_s1
                                )
  #flextable(final$Quantile)
  #paged_table(final$Sum_report)

paged_table(final$df_thetas)
paged_table(final$df_QOI)
paged_table(final$df_Decision)


```
